---
title: "DAFITI - OKR 2020 H1"
date: 2020-02-07
output:
  html_document:
    df_print: paged
    theme: cerulean
    highlight: zenburn
    fig_width: 8.5
    fig_height: 4
    code_folding: hide
---

```{r, eval=T,include=F,warning=F,message=F,echo=F}
library(tidyverse)
```

#### **Objective 1:** **Live the #ONEDFT culture where everyone feels responsible, belonging and empowered to revolutionize the way we work**

1. **KR 1- Reduce cancelation rate in 90 bps when comparing with the first half of 2019** `r emo::ji("voltage")` `r emo::ji("voltage")`

<button class="btn btn-primary btn-sm" data-toggle="collapse" data-target="#BlockName1"> KR 1 Follow Up </button>  
<div id="BlockName1" class="collapse"> 
```{r, eval=T,include=T,warning=F,message=F,echo=F}
mtcars %>% ggplot2::ggplot(aes(x = drat, y = mpg))+
  geom_point()
```
</div>

- [X] Criar trava de boleto para reduzir geração de boleto por boletadores. Trava deverá estar em produção até o final do semestre `r emo::ji("medal")` 

- [ ] Mapear cancelamentos por falta de estoque

2. **KR 2- Reduce cancelation rate in 90 bps when comparing with the first half of 2019** `r emo::ji("voltage")`

<button class="btn btn-primary btn-sm" data-toggle="collapse" data-target="#BlockName2"> KR 2 Follow Up </button>  
<div id="BlockName2" class="collapse"> 
```{r, eval=T,include=T,warning=F,message=F,echo=F}
mtcars %>% head()
```
</div>

#### **Objective 2:** **teste teste**

