---
title: "Monte Carlo Simulation Applied on Discounted Cash Flow"
author:
  - Francisco Piccolo

output:
  bookdown::pdf_document2:
    fig_caption: yes
    fig_width: 7.5
    fig_height: 6
    df_print: kable
    keep_tex: yes
    number_sections: no
    toc: no
    includes:
      in_header: header.tex
    
documentclass: article
# classoption: twocolumn, landscape, etc
papersize: a4
pagestyle: plain  # plain, empty, headings
geometry: margin=1in
linestretch: 1.2
fontfamily: mathpazo
fontsize: 11pt

# zotero: "My shared Zotero group"
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

linkcolor: RoyalBlue
urlcolor: RoyalBlue
links-as-notes: false
---

\normalsize

\vspace{1cm}
\hrule

My abstract.

\vspace{3mm}
\hrule
\vspace{5mm}

*Keywords*: My Keywords

\bleft
\newpage

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(knitr)

## Text results
opts_chunk$set(echo = F, warning = F, message = F, fig.align = "center")

## Code decoration
opts_chunk$set(tidy = T, comment = NA, highlight = T)

## Plots
opts_chunk$set(fig.path = "output/figures/")

```

# INTRODUCTION

Write your introduction here. You can cite bibliography like this [@Yan2011; @Sutherland2011] if you provide a `BibTeX` file with references. See http://rmarkdown.rstudio.com for more information. Or you could also use [knitcitations](https://cran.r-project.org/web/packages/knitcitations/index.html) or [RefManageR](https://cran.r-project.org/web/packages/RefManageR/index.html) to fetch bibliographic metadata automatically from the web. For example, citing a paper can be as easy as providing its DOI.

You can even specify the desired output format for your bibliography by including a style file for a specific journal (e.g. "ecology.csl"). Many different bibliography styles (CSL files) can be obtained at http://citationstyles.org/, https://github.com/citation-style-language/styles, or https://zotero.org/styles.

# METHODS

## Study Area

We worked in a **beautiful** place with lots of trees, like *Quercus suber* and *Laurus nobilis*.


## Data collection and analysis

```{r read_data, eval = FALSE}
dataset <- read.csv("mydata.csv")
```

We applied a linear model where

$$
y_{i} = \alpha + \beta*x_{i} 
$$

```{r model, eval = FALSE}
model <- lm(y ~ x)
```

# RESULTS

Trees in forest *A* grew taller than those in forest *B* (mean height: `r mean(25, 31, 28)` versus `r mean(13, 19, 16)` m). 

And many more cool results that get updated dynamically, e.g. see Table \@ref(tab:Table-mtcars) and Fig. \@ref(fig:scatterplot). Note Tables and Figures are cross-linked and numbered automatically. They could also appear in the middle of the document, not necessarily at the end.


# DISCUSSION

Discuss.

# CONCLUSIONS

Wrap up

\eleft

\clearpage


\listoftables


\newpage

\clearpage

\clearpage

\listoffigures

\newpage



